--- 
title: "Modeling the Spread of Norovirus on Commercial Cruise Voyages in R"
author: "Robert Sholl, Kelechi Igwe, Stephen Edache"
date: "`r Sys.Date()`"
site: bookdown::bookdown_site
documentclass: book
bibliography: [book.bib, packages.bib]
fontsize: 11
description: this is the tutorial for reproducing the code used in the STAT764 final project by Robert Sholl, Kelechi Igwe, and Stephen Edache.
link-citations: yes
---

```{r, include=FALSE}
options(tinytex.verbose = TRUE)
```


# About {-}

The following document details a step-by-step explanation for how to utilize R and data collected from the CDC Vessel Sanitation Program (VSP) to determine the rate of transmission of Norovirus on commercial cruise ships. 

This will include shallow overviews of the mathematical justifications behind why certain functions and data cleaning measures will taken, but for a more rigorous synopsis please refer to our article.

```{r include=FALSE}
# automatically create a bib database for R packages
knitr::write_bib(c(
  .packages(), 'bookdown', 'knitr', 'rmarkdown'
), 'packages.bib')
```
